---
title: "Projects"
---

There is a growing list of personal projects that I am excited to dive into as time goes on. A lot of them are focused on sports statistics or fun data analysis. Here I will provide a brief overview of what I have done, as well as links to the full body of work.

## [Runs Batted In Vs Total Opportunities](https://github.com/tdpotter58/rbis-per-opportunity)

```{r libraries, echo = FALSE, warning=FALSE, message=FALSE}
library(dplyr)
library(tidyr)
library(knitr)
library(readr)
library(ggplot2)
library(ggrepel)
library(stringr)
```

```{r rbi1, echo = FALSE, fig.width= 12, fig.height = 8}
theme_set(theme_bw())
rbis_full <- read.csv("data/rbis-full.csv")
rbis_full <- rbis_full %>%
  mutate(rbi_percentage = round(RBI / Chances, digits = 4))

rbis_2020 <- read.csv("data/rbis-2020.csv")
rbis_2020 <- rbis_2020 %>%
  mutate(rbi_percentage = round(RBI / Chances, digits = 4))

mean_2020 <- sum(rbis_2020[["RBI"]]) / sum(rbis_2020[["Chances"]])
rbis_2020 %>%
  filter(Chances > 4) %>%
  ggplot(aes(x = Chances, y = rbi_percentage, label = Player)) + 
  geom_hline(yintercept = mean_2020, linetype = "dashed", color = "red", size = 2) +
  geom_point(data = subset(rbis_2020, Chances >= 29 | (Chances > 4 & rbi_percentage > 0.65))) + 
  geom_text(data = subset(rbis_2020, Chances < 29 & (Chances > 4 & rbi_percentage <= 0.65)),size = 3) +
  geom_label_repel(data = subset(rbis_2020, Chances >= 29 | (Chances > 4 & rbi_percentage > 0.65)), aes(fill = RBI),size = 3) +
  scale_fill_gradient(low = 'blue', high = 'red') +
  ggtitle("RBI Earned with Runner at Third Base, 2020") +
  xlab("At-Bats with a Runner at Third Base")+
  ylab("Proportion of Runners Batted In") +
  theme(plot.title = element_text(hjust = 0.5)) +
  labs(caption = "League Average = 0.376")
```
A look at how often each MLB player came to the plate with a runner at third base, and how often they were able to score said runner. I wanted to see what players were the most efficient in this situation, and what players just had better chances to inflate their RBI numbers. Data was scraped from Baseball-Reference for the 2016-2020 MLB seasons.

```{r rbi, echo = FALSE, fig.align='center', fig.width= 15, fig.height = 10}


kable(head(rbis_full, n = 15))

mean_full <- sum(rbis_full[["RBI"]]) / sum(rbis_full[["Chances"]])
rbis_full %>%
  filter(Chances > 20) %>%
  ggplot(aes(x = Chances, y = rbi_percentage, label = Player)) + 
  geom_hline(yintercept = mean_full, linetype = "dashed", color = "red", size = 2) +
  geom_text() +
  scale_fill_gradient(low = 'blue', high = 'red') +
  ggtitle("RBI Earned with Runner at Third Base, 2016-2020") +
  xlab("At-Bats with a Runner at Third Base")+
  ylab("Proportion of Runners Batted In") +
  theme(plot.title = element_text(hjust = 0.5)) +
  labs(caption = "League Average = 0.384")
```

The full analysis, as well as the code and datasets, can be found in my GitHub.

## [Pitching Extremes](https://github.com/tdpotter58/pitching-extremes)

```{r pitching, echo = FALSE, message=FALSE, fig.align='center', fig.width= 10, fig.height = 6}
pitchers <- read_csv("data/pitcher-data.csv")
pitchers <- pitchers %>%
  select(-Pit, -Str) %>%
  drop_na()
pitchers <- pitchers %>%
  mutate(Year_ch = as.character(pitchers$Year))
         
for (i in pitchers$Rk){
  name <- str_split(pitchers[i,2], "-")[[1]]
  
  pitchers[i,2] <- str_c(name[1],pitchers[i,48], sep = " ")
}
pitchers <- pitchers %>%
  mutate(Player.f = as.factor(pitchers$Player))
pitchers$Player.f <- factor(pitchers$Player.f, levels = pitchers$Player.f[order(pitchers$`W-L%`)])

theme_set(theme_bw())
pitchers %>%
  filter(`W-L%` < 0.5) %>%
  ggplot(aes(x = ERA, y = `W-L%`, label = Player)) + 
  geom_point(size = 2) +
  geom_label_repel(data = subset(pitchers, `W-L%` < 0.34), fill = 'cyan') +
  geom_label_repel(data = subset(pitchers, ERA < 2.25), fill = 'coral') +
  ylab("Winning Percentage") +
  ggtitle("Every Qualified Season Since 1968 with an ERA Below 3.00 and More Losses than Wins") +
  labs(caption = "Data from Baseball-Reference") +
  theme(plot.title = element_text(hjust = 0.5)) 
```

Sometimes in baseball, a pitcher can have a tremendous season throwing the ball and get no run support from his offense. In other cases the reverse happens, and a mediocre pitcher reaps the benefits of playing on a good team. This is a quick visualization of seasons where a pitcher was credited with more losses or wins than they deserved. 

## [Other Projects](https://github.com/tdpotter58)

Looking for more? Here's a mixture of classwork and other personal projects that are incomplete/in the works. More projects will be added above when I finally get around to putting them together.
